An Analysis of Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters

### 2010-10-24 doc/isbi/main.tex
added another chart to main.tex
getting close to the end

\section{Results}
\label{sec:Results}
A primary advantage of the particle filter is the that calculates
a complete joint distribution of the parameters and states. This 
allows for advanced analyses not available to many other methods.
Thus, the correlation matrix of parameters from a single 40 minute
simulated fMRI regression are shown in. %\autoref{
\begin{table}[t]
\ninept
  \centering
\begin{tabular}{|c | c  c  c  c  c  c  |}
\hline
  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$  \\
\hline
$\alpha$                      & 0.880& & & & & \\
\rowcolor[gray]{.8} $E_0$     & -0.766& -0.523& & & & \\
$V_0$                         & 0.624& 0.424& -0.796& & & \\
\rowcolor[gray]{.8} $\tau_s$  & 0.620& 0.295& -0.748& 0.344& & \\
$\tau_f$                      & 0.000& -0.397& -0.431& 0.196& 0.699& \\
\rowcolor[gray]{.8} $\epsilon$& 0.616& 0.656& -0.641& 0.285& 0.446& -0.097\\
\hline
\end{tabular}
  \caption{Correlation Matrix of Posterior Distribution for 40 minute simulated BOLD signal.}
\label{tab:long_corr}
\end{table}[htb]
\normalsize

For POSSUM simulations, \autoref{fig:PossumResult} shows the
the mutual information between the estimated timeseries and the 
preprocessed signal. The estimate was generated from the 
mean of the particles' parameters at the end of the run.

@@

\section{Discussion}
\label{sec:Conclusion}

\bibliographystyle{IEEEbib}
\bibliography{/home/micahc/library}

>>>>

\section{Results}
\label{sec:Results}
A primary advantage of the particle filter is that it estimates
a complete joint distribution of the parameters and states. This 
allows for more detailed analyses not available to many other methods.
In particular it is possible to calculate the correlation of the
parameters in the posterior distribution. The correlation matrix
for the single 40 minute simulation is shown in \autoref{tab:long_corr}.
\begin{table}[t]
\ninept
  \centering
\begin{tabular}{|c | c  c  c  c  c  c  |}
\hline
  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$  \\
\hline
$\alpha$                      & 0.880& & & & & \\
\rowcolor[gray]{.8} $E_0$     & -0.766& -0.523& & & & \\
$V_0$                         & 0.624& 0.424& -0.796& & & \\
\rowcolor[gray]{.8} $\tau_s$  & 0.620& 0.295& -0.748& 0.344& & \\
$\tau_f$                      & 0.000& -0.397& -0.431& 0.196& 0.699& \\
\rowcolor[gray]{.8} $\epsilon$& 0.616& 0.656& -0.641& 0.285& 0.446& -0.097\\
\hline
\end{tabular}
  \caption{Correlation Matrix of Posterior Distribution for 40 minute simulated BOLD signal.}
\label{tab:long_corr}
\end{table}[htb]

For POSSUM simulations, \autoref{fig:PossumResult} shows the
the mutual information between the estimated timeseries and the 
preprocessed signal. Each estimate was generated from the 
mean of the particles' parameters at the end of the run. 
Additionally the mean and variance of the parameter estimates
(mean) across the spatial regions of \autoref{fig:PossumResult} are
given in \autoref{tab:beforeafter}.

@@

\begin{table}[t]
\ninept
  \centering
\begin{tabular}{| c | c | c | c | c | c |}
\hline
& Sim Value & \multicolumn{2}{|c|}{Prior} &\multicolumn{2}{|c|}{Posterior} \\
\hline
        &       &$\mu$  & $\sigma$& $\mu$& $\sigma$\\
\hline
$\tau_0  $& 1.45  & .98   & .25     & 1.07319 & 0.297749  \\
\rowcolor[gray]{.8} 
$\alpha  $&0.3    & .33   & .045    & 0.317955& 0.07311  \\
$E_0     $& 0.47  & .34   & .03     & 0.364195& 0.0457019  \\
\rowcolor[gray]{.8} 
$V_0     $& 0.044 & .04   & .03     & 0.110542& 0.0500039  \\
$\tau_s  $& 1.94  & 1.54  & .25     & 1.86785 & 0.390455  \\
\rowcolor[gray]{.8} 
$\tau_f  $& 1.99  & 2.46  & .25     & 2.50166 & 0.348326  \\
$\epsilon$& 1.8   & .7    & .6      & 1.58716 & 1.14536  \\
\hline
\end{tabular}
\caption{Comparison of the true value (left), the prior distribution
used to generate the initial particles (center two columns) and the 
average/variance of the estimates across simulated voxels. Note that
each voxel used the same underlying parameters (with differences only
in absolute value and noise realization), and all particle filters
were initialized with the same prior.}
\label{tab:beforeafter}
\end{table}[htb]

\section{Discussion}
\label{sec:Conclusion}
The data shows substantial correlation present in the parameter 
estimates for the 40 minute simulated data. Correlation of 
constants is a definite sign of the model being under-determined
and thus not observable. This is in spite of low noise, and the 
particle filter's solution converging to a Root-Mean-Squared error
of less than $0.002$.

On a larger scale, the particle filter also performed well, clearly
locating regions of activation in spite of Signal-to-Noise Ratio 
well below 6dB. In spite of this, the final parameter estimates 
varied more than the initial (prior) distribution (\autoref{tab:beforeafter}).
This effect is slightly deceiving because the distributions
were far from Gaussian, and many were bi-modal. It is 
interesting, however, that such a wide range of solutions
could give good estimates for essentially one BOLD signal.

From the results, it is clear that the BOLD model as presented
by Friston et al. is not observable, and thus parameter estimates
suffer from very large variance \cite{Friston2000}. As a result, parameter estimates
in that work are likely the result in the under-determination of
the model rather than underlying differences in parameters. 
There are two possible ways to deal with this problem. First, as
specified in Deneux et al., certain parameters could be locked into
place \cite{Deneux2006}. This solution is relatively easy, and would
also reduce computation time. True in-vivo parameter estimation
would help to determine proper values for parameters that are to be
kept constant, although it is a difficult proposition. The other
possible method is adding another measurement point. Technology exists
to simultaneously measure fMRI and \ac{CBF}. This additional
measurement, which is slightly closer to original stimulus, could
make the Balloon Model observable.

\small
\bibliographystyle{IEEEbib}
\bibliography{/home/micahc/library}
